---
title: "Peer Assesment 1"
author: "Gutxi"
date: "Thursday, July 16, 2015"
output: html_document
---

```{r}
library(ggplot2)

activity <- read.csv("~/data/activity.csv")
```

**Number of steps by day**
```{r}
stepsByDay <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)

```

Histogram
```{r}
qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency')
```

Mean and median of the total steps by day
```{r}
stepsByDayMean <- mean(stepsByDay)
stepsByDayMean 
stepsByDayMedian <- median(stepsByDay)
stepsByDayMedian
```
**Average daily activity pattern**
```{r}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=activity$steps), by=list(interval=activity$interval), FUN=mean, na.rm=TRUE)
```

Time series plot
```{r}
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```

5-minute interval with maximum number of steps
```{r}
mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
timeMostSteps
```
**Imputing missing values**
Calculate and report the total number of missing values in the dataset
```{r}
MissingValues <- length(which(is.na(activity$steps)))
MissingValues
```

New dataset that is equal to the original dataset but with the missing data filled in
```{r}
na_pos <- which(is.na(activity$steps))
mean_vec <- rep(mean(activity$steps, na.rm=TRUE), times=length(na_pos))
activity[na_pos, "steps"] <- mean_vec
head(activity)
```
histogram of the total number of steps taken each day
```{r}
TotalstepsByDay <- tapply(activity$steps, activity$date, sum)
qplot(TotalstepsByDay, xlab='Total steps per day', ylab='Frequency', binwidth=500)
```

Mean and median
```{r}
TotalstepsByDayMean<- mean(TotalstepsByDay)
TotalstepsByDayMean
TotalstepsByDayMedian<- median(TotalstepsByDay)
TotalstepsByDayMedian
```

**Differences in activity patterns between weekdays and weekends**
new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day
```{r}
activity$dateType <-  ifelse(as.POSIXlt(activity$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

Time series plot
```{r}
averagedActivity <- aggregate(steps ~ interval + dateType, data=activity, mean)
ggplot(averagedActivity, aes(interval, steps)) + geom_line() + facet_grid(dateType ~ .) + xlab("5-minute interval") + ylab("average number of steps")
```


